---
title: "Trip 1"
output: html_document
date: "2024-02-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load in Libraries

```{r, echo=FALSE}


library(sf)
library(raster)
library(readxl)
library(dplyr)
library(ggplot2)
library(osmdata)
library(leaflet)
library(raster)
library(gstat)
library(akima)


```

Load in Data and Filter out uneeded columns

```{r, echo=FALSE}
#read in data
ds <- read_excel("Trip_1_Data.xlsx")

#Filter columns
ds <- subset(ds, select = c("Delta C (ug/m^3)", "Latitude", "Longitude", "3 Sec Moving Avg (Delta C ug/m^3)"))


#Rename columns
#Switching Lat and Long data 
colnames(ds)[colnames(ds) == "Delta C (ug/m^3)"] <- "Delta_C"
colnames(ds)[colnames(ds) == "3 Sec Moving Avg (Delta C ug/m^3)"] <- "Delta_C_avg"
#colnames(ds)[colnames(ds) == "Longitude"] <- "latitude"
#colnames(ds)[colnames(ds) == "Latitude"] <- "longitude"
#colnames(ds)[colnames(ds) == "longitude"] <- "Longitude"
#colnames(ds)[colnames(ds) == "latitude"] <- "Latitude"

```

NOTE: MAKE THE ZERO VALUE TRANSPARENT IF POSSIBLE


```{r, echo=FALSE}
# Load necessary libraries
library(sp)
library(raster)
library(ggplot2)
library(gstat)

# Converts the coordinates in the data set to spatial data 
coordinates(ds) <- ~Longitude+Latitude 

# Define the extent and create a raster grid
# Make sure the x.range uses Longitude and y.range uses Latitude
x.range <- range(ds@coords[,1])  # Computes Longitude range (first column)
y.range <- range(ds@coords[,2])  # Computes Latitude range (second column)

#Defines the extent of the raster grid using the min and max Lat and Lon values
#Converts the 30m x 30m resolution to degrees 
#111319.9 is an approximation of how many meters there are in a degree of latitude or longitude at the equator: Earth circumference / 360 degrees
extent_raster <- raster(extent(x.range[1], x.range[2], y.range[1], y.range[2]), res=10/111319.9)

# Create a set of points from the raster extent for interpolation
# Converts it to a data frame 
# Converts the df to spatial data/coordinates
grid_points <- rasterToPoints(extent_raster)
grid_points_df <- data.frame(grid_points)
coordinates(grid_points_df) <- ~x+y  

# Perform IDW interpolation using the grid points for Delta_C
# formula ~ 1 meanse it is only considering Delta_C
# newdata is where we want to interpolate the data
# IDP = Inverse Distance Power which controls how the weigth decreases with distance. A value of 2.0 means that the influence of a point on the interpolation decreases as the square of the distance. 
idw_output <- idw(formula = Delta_C_avg ~ 1, locations = ds, newdata = grid_points_df, idp = 2.0)

# Convert IDW output to SpatialPixelsDataFrame and then to raster for plotting
gridded(idw_output) <- TRUE # Converts the idw_output which is a SpatialPointsDataFrame, to a SpatialPixelDataFrame for plotting
raster_interp <- raster(idw_output) # Converts it to the gridded data to a raster format

# Set values less than X to NA to omit non-needed raster vals
#raster_interp[raster_interp < 0.5] <- NA

# Defines a color palette for the heatmap
color_palette <- colorRampPalette(c("darkslategray1", "darkturquoise", "blue", "coral", "coral1", "coral2", "coral3", "coral4", "red", "darkred"))


# Plot the heatmap
# zlim defines the range 
plot(raster_interp, main="Heat Map of Delta_C over Sechelt, BC", col=color_palette(100), xlab="Longitude", ylab="Latitude", zlim=c(min(ds$Delta_C_avg, na.rm=TRUE), max(ds$Delta_C_avg, na.rm=TRUE)))

```
Raster grid plotted over a map of Sechelt, BC

```{r}
# Note just load them in everytime for simplicity when editing different code chunks
# Load the necessary libraries
library(raster)   # For handling raster data
library(leaflet)  # For interactive maps
library(sp)       # For spatial data classes and methods

# 'raster_interp' is our raster object
# Uses an if statement with an "or" (||) condition 
# proj4string(raster_interp) gets the coordinate reference system (CRS) information of raster_interp in a text format called proj4string
# CRS() analyzes the proj4string and determines what it is saying. If it is not correct, it will return an NA val which is.na() will catch 
# Second condition detemrines if it an empty string
# crs(raster_interp) is about assigning a new CRS to raster_interp
# CRS("+proj=longlat +datum=WGS84 +no_defs"): we're defining what the new CRS should be: +proj=longlat: This tells us the projection system is based on longitude and latitude. 
# +datum=WGS84: This specifies the datum (a reference model of the Earth) used for the projection. WGS84 is a widely used global datum. 
# +no_defs: means no additional definitions are needed for this CRS.

if (is.na(CRS(proj4string(raster_interp))) || CRS(proj4string(raster_interp)) == "") {
  crs(raster_interp) <- CRS("+proj=longlat +datum=WGS84 +no_defs")
}

# Plot the raster on a Leaflet map centered on Sechelt, BC
leaflet() %>%
  addTiles() %>%
  addProviderTiles("Thunderforest.MobileAtlas") %>%
  setView(lng = -123.7536, lat = 49.4742, zoom = 13) %>%
  addRasterImage(raster_interp, opacity = 0.6, colors = colorRampPalette(c("darkslategray1", "darkturquoise", "blue", "coral", "coral1", "coral2", "coral3", "coral4", "red", "darkred"))(100))

```
Now Doing Raster Map For Mobile Home Park Thing
```{r}
library(leaflet)

#read in data
ds <- read_excel("Trip_1_Data.xlsx")

#Filter columns
ds <- subset(ds, select = c("Delta C (ug/m^3)", "Latitude", "Longitude", "3 Sec Moving Avg (Delta C ug/m^3)"))


#Rename columns
#Switching Lat and Long data 
colnames(ds)[colnames(ds) == "Delta C (ug/m^3)"] <- "Delta_C"
colnames(ds)[colnames(ds) == "3 Sec Moving Avg (Delta C ug/m^3)"] <- "Delta_C_avg"
#colnames(ds)[colnames(ds) == "Longitude"] <- "latitude"
#colnames(ds)[colnames(ds) == "Latitude"] <- "longitude"
#colnames(ds)[colnames(ds) == "longitude"] <- "Longitude"
#colnames(ds)[colnames(ds) == "latitude"] <- "Latitude"

#Create the heatmap with route
HeatRoute <- leaflet() %>%
  addTiles() %>%
  addProviderTiles("Thunderforest.MobileAtlas")


# Add the raster image with the specified color palette
HeatRoute <- HeatRoute %>% addPolylines(data = ds, ~Longitude, ~Latitude, color = "red", weight = 2, opacity = 0.4) %>%
  addRasterImage(raster_interp, opacity = 0.4, 
                 colors = colorRampPalette(c("darkslategray1", "darkturquoise", "blue", "coral", "coral1", "coral2", "coral3", "coral4", "red", "darkred"))(100)) %>%
  setView(lng = mean(ds$Longitude), lat = mean(ds$Latitude), zoom = 14)

# Print the map
HeatRoute
```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```


This code should work, and I cannoy figure out why the heatmap doesn't show up.
```{r}


```


This just creates a basic heat map with arbitrary points

I just put it here for reference
```{r}


```



